---
title: "Exploratory Data Analysis"
author: "Michael Larbie, Sarah Piper, Nikhita Gopisetty"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

##Import Packagaes
```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(viridis)
library(ggpubr)
library(FactoMineR)
library(factoextra)
```

## Load Data
```{r}
data <- read_csv('data/adult21.csv')
```

## Native Born Compared to Non-Native Born Covid Diagnosis
```{r}
data$NATUSBORN_A <- as.factor(data$NATUSBORN_A)
data$CVDDIAG_A <- as.factor(data$CVDDIAG_A)

data %>% count(NATUSBORN_A,CVDDIAG_A) %>%
ggplot(aes(fill = CVDDIAG_A, x = NATUSBORN_A, y =n)) +
geom_bar(position = "fill", stat = "identity") + 
  labs(title = "Native v Non-Native Born", x = "Origin of Birth",
                  y = "Number of Respondents")
```

## Region Positive Covid Diagnosis
```{r}
data$REGION <- as.factor(data$REGION)


data %>% count(REGION,CVDDIAG_A) %>% 
  ggplot(aes(fill = REGION, x = CVDDIAG_A, y = n)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(title = "Region Positive Covid Diagnosis", x = "Covid Diagnosis",
       y = "Number of Respondents")

data %>% count(NATUSBORN_A,REGION) %>% 
  ggplot(aes(fill = NATUSBORN_A, x = REGION, y = n)) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(title = "Regional Nativity Status", x = "Region",
       y = "Number of Respondents")
```

## Diseases and Covid Diagnosis
```{r}
data$HYPEV_A <- as.factor(data$HYPEV_A)
data$CHLEV_A <- as.factor(data$CHLEV_A)
data$ASEV_A <- as.factor(data$ASEV_A)
data$DIBEV_A <- as.factor(data$DIBEV_A)

hypertension <- data %>% count(HYPEV_A,CVDDIAG_A, NATUSBORN_A) %>% 
  ggplot(aes(fill = HYPEV_A, x = NATUSBORN_A, y = n)) +
  geom_bar(position = "fill", stat = "identity") +
  facet_wrap(~ CVDDIAG_A) +
  labs(title = "Hypertension and Covid Diagnosis", x = "Nativity",
       y = "Fraction of Respondents")

cholesterol <- data %>% count(CHLEV_A,CVDDIAG_A, NATUSBORN_A) %>% 
  ggplot(aes(fill = CHLEV_A, x = NATUSBORN_A, y = n)) +
  geom_bar(position = "fill", stat = "identity") +
  facet_wrap(~ CVDDIAG_A) +
  labs(title = "Cholesterol and Covid Diagnosis", x = "Nativity",
       y = "Fraciton of Respondents")


asthma <- data %>% count(ASEV_A,CVDDIAG_A, NATUSBORN_A) %>% 
  ggplot(aes(fill = ASEV_A, x = NATUSBORN_A, y = n)) +
  geom_bar(position = "fill", stat = "identity") +
  facet_wrap(~ CVDDIAG_A)+
  labs(title = "Asthma and Covid Diagnosis", x = "Nativity",
       y = "Fraction of Respondents")


diabetes <- data %>% count(DIBEV_A,CVDDIAG_A, NATUSBORN_A) %>% 
  ggplot(aes(fill = DIBEV_A, x = NATUSBORN_A, y = n)) +
  geom_bar(position = "fill", stat = "identity") +
  facet_wrap(~ CVDDIAG_A)+
  labs(title = "Diabetes and Covid Diagnosis", x = "Nativity",
       y = "Fraction of Respondents")

ggarrange(hypertension, cholesterol, asthma, diabetes, nrow = 2, ncol = 2 )
```

## Social Determinants, Nativity, COVID
```{r}
data$MEDNG12M_A <- as.factor(data$MEDNG12M_A)
data$PAYWORRY_A <- as.factor(data$PAYWORRY_A)
data$FDSRUNOUT_A <- as.factor(data$FDSRUNOUT_A)

data %>% count(MEDNG12M_A,NATUSBORN_A, CVDDIAG_A) %>% 
ggplot(aes(fill = MEDNG12M_A, x = NATUSBORN_A, y = n)) +
geom_bar(position = "fill", stat = "identity") +
facet_wrap(~CVDDIAG_A) +
labs(title = "Access to Care and Covid Diagnosis", x = "Access to Care",
       y = "Number of Respondents")

data %>% count(PAYWORRY_A, NATUSBORN_A, CVDDIAG_A) %>% 
ggplot(aes(fill = PAYWORRY_A, x = NATUSBORN_A, y = n)) +
geom_bar(position = "fill", stat = "identity") +
facet_wrap(~CVDDIAG_A) +
labs(title = "Ability to Pay for Healthcare and Covid Diagnosis", 
     x = "Nativity",
       y = "Number of Respondents")

data %>% count(FDSRUNOUT_A,NATUSBORN_A, CVDDIAG_A) %>% 
ggplot(aes(fill = FDSRUNOUT_A, x = NATUSBORN_A, y = n)) +
geom_bar(position = "fill", stat = "identity") +
facet_wrap(~CVDDIAG_A) +
labs(title = "Food Security and Nativity", x = "Nativity",
       y = "Number of Respondents")
```

## Race, Nativity, Covid
```{r}
data$RACEALLP_A <- as.factor(data$RACEALLP_A)

data %>% mutate(race = case_when(RACEALLP_A == 1 ~ "White",
                                RACEALLP_A == 2 ~ "Black",
                                RACEALLP_A == 3 ~ "Asian",
                                RACEALLP_A == 8 ~ "Latino")) %>% 
mutate(immigrant = case_when(NATUSBORN_A == 1 ~ "Native",
                                NATUSBORN_A == 2 ~ "Immigrant"))%>%
mutate(covid = case_when(CVDDIAG_A == 1 ~ "Positive",
                             CVDDIAG_A == 2 ~ "Negative")) %>%
drop_na(race, immigrant, covid) %>%
count(race, immigrant, covid) %>%
ggplot(aes(fill = covid, x = race, y = n)) +
geom_bar(position = "fill", stat = "identity") +
facet_wrap(~ immigrant)+
labs(title = "Relationship Between Race, Immigration and Covid Diagnosis", 
     x = "Race",
     y = "Fraction of Respondents")
```
In our EDA we wanted to find trends mainly between three categorical groups that
being immigration, COVID-19 diagnosis and a third variable. In doing our EDA as 
can be seen in figure 1 across all racial groups, those of non US birth or 
immigrants had a higher percentage of their sample population being positive for
COVID compared to their, US born counter parts.This same trend was found when 
comparing social determinants, immigration status and COVID diagnosis. For 
access to care, difficulty paying for healthcare and food security, immigrants 
always had a higher percentage of their population with COVID, who felt at 
higher risk of not getting these basic necessities compared to their US born 
counterparts. Diabetes and high cholesterol followed this trend as well, but 
with just slightly higher percentages of their population having the disease
prior to having COVID compared to their US born counterparts.
This trend could not be seen with different pre-conditions as the 
third variable, these diseases being asthma and hypertension. The immigrant 
population had the same or sometimes lower percentages of their population with 
the disease who also had COVID. In doing this EDA we discovered that immigrant
status does have an affect on COVID diagnosis and does influence all social
determinant variables used in the EDA. While it seems to have little affect on 
pre conditions. Our group will need to reconsider the pre-condition variables 
used in our data set and see if there are other preconditions in the data set
which have more of a relationship with nativity and COVID. We will have to
do some pre-processing to get rid of results that are not needed for our 
analysis such as for option in the survey for people whole refused to give up 
some information or did not know the information. There were no NA's in the
variables we used so that is not something we need to worry about.

## MCA
```{r}
data2 <- data %>% select(REGION, CVDDIAG_A, NATUSBORN_A, RACEALLP_A,MEDNG12M_A,
                         PAYWORRY_A, FDSRUNOUT_A, DIBEV_A, ASEV_A, CHLEV_A,
                         HYPEV_A) %>% 
  filter(NATUSBORN_A == 1 | NATUSBORN_A == 2, MEDNG12M_A ==1 | MEDNG12M_A == 2,
         PAYWORRY_A == 1 | PAYWORRY_A == 2 | PAYWORRY_A == 3, 
         FDSRUNOUT_A == 1 | FDSRUNOUT_A == 2 | FDSRUNOUT_A == 3,
         CVDDIAG_A == 1 | CVDDIAG_A == 2, DIBEV_A == 1 | DIBEV_A == 2,
         ASEV_A == 1 | ASEV_A == 2, CHLEV_A == 1 | CHLEV_A == 2, 
         HYPEV_A == 1 | HYPEV_A == 2)
  
res <- MCA(data2)
fviz_mca_var(res, 
            repel = TRUE, # Avoid text overlapping (slow)
            ggtheme = theme_minimal())

```

Our group also conducted a Multiple Correspondence Analysis on the different 
variables we intended to use in our research to find which variables had the 
most similar profiles in the sampled population used. By doing the MCA we 
wanted to find which variables would be useful in helping to understand our 
hypothesis and which variables we selected had no significant impact. When 
interpreting MCA variable categories closer together tend to share similar 
responses among the population. From conducting our MCA and as can be seen in 
figure 2 it was found that the negative COVID diagnosis variable or CVDDIAG_A_2 
shared similar profiles to those without diabetes, asthma, and high cholesterol.
The negative COVID diagnosis variable also shared similar profiles to those 
who did not worry about food security, where able to get medical care
regardless of cost and where somewhat worried about being able to pay medical
bills. The variable positive COVID diagnosis or CVDDIAG_A_1 did not have any 
close relationships with other variables. From our MCA we concluded that the 
lack of prior health risks was more associated with not getting COVID, while on
the other hand there did not seem to be any correlation between having these 
prior health risks and being more likely to get COVID. The same sentiment can be
used on social determinants those with a lower risk of having food security,
access to care and ability to afford health share a similar profile to a 
negative COVID diagnosis, while those at higher risk do not share that same
profile with a positive COVID diagnosis. Looking at the MCA there was a much
closer relationship between those born in the US (NATUSBORN_A_1) and a negative
COVID diagnosis, than compared to those born outside the US and a positive 
COVID diagnosis.